FreeAdaptation20160417
===================

This is a free adaptation log, the mice had 2 days of free adaptation with no cue, door is always open. and on the 17th we replaced it to a free adaptation: between 19:00-24:00 the door opens and a light cue apears for 5 sec, when the light disappears the door is closed and it wouldn't open until a new visit will occour. between 24:00-19:00 the door is closed and when the mice nosepoke it they get a 15kHz sound for negative reinforcement.

This R markdown requires ggplot2 package and will install it if it doesn't exist.  
```{r, echo=FALSE}
#check if ggplot exists and install it if not
list.of.packages <- c("ggplot2")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

It read the "animals.txt", "visits.txt" and "nosepokes.txt" files,    
```{r, echo = FALSE}
#Read the intellicage data, nosepokes, visits, and animals
sessionPath <- "Sessions/2016-04-17 16.40.44/"
visits <- read.csv(paste0(sessionPath,"IntelliCage/Visits.txt"), sep = "\t")
nosepokes <- read.csv(paste0(sessionPath,"IntelliCage/Nosepokes.txt"), sep = "\t")
animals <- read.csv(paste0(sessionPath,"/Animals.txt"), sep = "\t")
```
and assigns animal name for each event as a factor(visit and nosepoke).  

```{r, echo = FALSE}
#Replace AnimalTag with its AnimalName
visits$AnimalName <- lapply(visits$AnimalTag,
                                function(x){
                                            animals$AnimalName[animals$AnimalTag == x ]})
visits <- transform(visits, AnimalName = unlist(AnimalName))

#Assigning an animal name to each visit id in the nosepokes data set
nosepokes$AnimalName <- lapply(nosepokes$VisitID,
                               function(x){
                                       visits$AnimalName[visits$VisitID == x ]})
nosepokes <- transform(nosepokes, AnimalName = unlist(AnimalName))

```

This is the visit count with the appropriate plot  
```{r, echo = FALSE}
#Plots visit count per animal
library(ggplot2)
visitsCountPlot <- ggplot(visits, aes(AnimalName)) +
                   geom_bar() +
                   xlab("") +
                   ggtitle("visits Count")
print(table(visits$AnimalName))
plot(visitsCountPlot)
```

Visits over time for each animal   
```{r, fig.width=10, fig.height=4, echo = FALSE}
#plot visits over time using the ggplot2 package
visitPoints <- geom_point(aes(x = as.POSIXct(visits$Start),
                              y = visits$AnimalName,
                              colour = visits$AnimalName),
                              alpha = 0.3)
visitPointsPlot <- ggplot() +
                   visitPoints +
                   xlab("") +
                   ylab("") +
                   ggtitle("Visits") +
                   theme(legend.title=element_blank())
plot(visitPointsPlot)
```

Visit count across absoulute time  
```{r, echo = FALSE}
#Count visits per hour for each animal
animalDat <- split(visits, visits$AnimalName)
visitHourAnimal <- lapply(animalDat, 
                          function(x){vapply(split(1:nrow(x),
                                            format(as.POSIXlt(x$Start),
                                                   "%Y-%m-%d %H:00:00",
                                                    tz='UTC')),
                                            length,
                                            0)})
visitHourAnimal <- unlist(visitHourAnimal)
visitHourAnimalDF <- read.table(text = names(visitHourAnimal), sep = ".")
visitHourAnimalDF$VisitCount <- as.numeric(visitHourAnimal)
names(visitHourAnimalDF) <- c("AnimalName", "Hour", "VisitCount")
visitHourAnimalDF$Hour <- as.POSIXct(visitHourAnimalDF$Hour)
```

```{r, echo = FALSE, fig.width = 12, fig.height = 6}
visitLines <- ggplot(visitHourAnimalDF,
                     aes(Hour, VisitCount, colour = AnimalName))
visitCountPlot <- visitLines + geom_line()
print(visitCountPlot)
```

```{r, echo = FALSE}
#create a data frame for visit per hour
hours <- vapply(split(1:nrow(visits),
                      format(as.POSIXlt(visits$Start),"%Y-%m-%d%H:00:00",tz='UTC')),
                      length,
                      0)
visitsPerHour <- as.numeric(hours)
hour <- as.POSIXct(names(hours))
visitCount <- data.frame(hour,visitsPerHour)
#Plot a line plot for the visitCount data frame
visitCountLine <- ggplot(visitCount, aes(hour, visitsPerHour)) +
                  geom_line() +
                  ylab("Count") +
                  xlab("") +
                  ggtitle("Visit count per hour")
```

This is the hour with the maximun number of visits  
```{r, echo = FALSE}
print(visitCount$hour[which.max(visitCount$visitsPerHour)])
```


  

