Task_20160522-20160523
=======================

Upon a visit a random delay timer of {1.5s, 2.5, 3.5} starts then a light cue appears for 0.9s within the time of the light cue it is possible to open the door with a nosepoke, otherwise it counts as a time error{with no negative reinforcement sound}.

On the 22th Mice {2,5,9} recived 10uL per gram of saline(i.e if a mouse weighs 25g it recived 250uL of saline) through intraperitoneal injection. Mice {4,6,13} recived 10mg/Kg CNO.

On the 23th Mice {4,6,13} recived 10uL per gram of saline(i.e if a mouse weighs 25g it recived 250uL of saline) through intraperitoneal injection. Mice {2,5,9} recived 10mg/Kg CNO.

This R markdown requires ggplot2 package and will install it if it doesn't exist.  
```{r, echo=FALSE}
#check if ggplot exists and install it if not
list.of.packages <- c("ggplot2")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
library(ggplot2)
library(dplyr)
```

It read the "animals.txt", "visits.txt" and "nosepokes.txt" files, removes all non-first nosepokes    
```{r, echo = FALSE}
#Read the intellicage data, nosepokes, visits, and animals
sessionPath1 <- "Sessions/2016-05-22 11.14.26/"
sessionPath2 <- "Sessions/2016-05-23 18.29.11/"
visits1 <- read.csv(paste0(sessionPath1,"IntelliCage/Visits.txt"), sep = "\t")
nosepokes1 <- read.csv(paste0(sessionPath1,"IntelliCage/Nosepokes.txt"), sep = "\t")
visits2 <- read.csv(paste0(sessionPath2,"IntelliCage/Visits.txt"), sep = "\t")
nosepokes2 <- read.csv(paste0(sessionPath2,"IntelliCage/Nosepokes.txt"), sep = "\t")
source("bindVisits.R")
source("bindNosepokes.R")
visits <- bindVisits(visits1, visits2)
nosepokes <- bindNosepokes(visits1, nosepokes1, nosepokes2)


#Erase the cage TimeError count, its malfunctioning in some way.
nosepokes$TimeError <- 0
#remove the nosepokes from the right side
nosepokes <- nosepokes[!nosepokes$SideError,]
#Keep only the first nosepoke in each visit
nosepokes <- nosepokes[!duplicated(nosepokes$VisitID),]

animals <- read.csv(paste0(sessionPath1,"/Animals.txt"), sep = "\t")
```
and assigns animal name for each event as a factor(visit and nosepoke).  

```{r, echo = FALSE}
#Replace AnimalTag with its AnimalName
visits$AnimalName <- lapply(visits$AnimalTag,
                                function(x){
                                            animals$AnimalName[animals$AnimalTag == x ]})
visits <- transform(visits, AnimalName = unlist(AnimalName))

#Assigning an animal name to each visit id in the nosepokes data set
nosepokes$AnimalName <- lapply(nosepokes$VisitID,
                               function(x){
                                       visits$AnimalName[visits$VisitID  
                                                         == x ]})
nosepokes <- transform(nosepokes, AnimalName = unlist(AnimalName))

```

This is the visit count with the appropriate plot  
```{r, echo = FALSE}
#Plots visit count per animal
visitsCountPlot <- ggplot(visits, aes(AnimalName)) +
                   geom_bar() +
                   xlab("") +
                   ggtitle("visits Count")
print(table(visits$AnimalName))
plot(visitsCountPlot)
```

Visits over time for each animal   
```{r, fig.width=10, fig.height=4, echo = FALSE}
#plot visits over time using the ggplot2 package
visitPoints <- geom_point(aes(x = as.POSIXct(visits$Start),
                              y = visits$AnimalName,
                              colour = visits$AnimalName),
                              alpha = 0.3)
visitPointsPlot <- ggplot() +
                   visitPoints +
                   xlab("") +
                   ylab("") +
                   ggtitle("Visits") +
                   theme(legend.title=element_blank())
plot(visitPointsPlot)
```

Visit count across absoulute time  
```{r, echo = FALSE}
#Count visits per hour for each animal
animalDat <- split(visits, visits$AnimalName)
visitHourAnimal <- lapply(animalDat, 
                          function(x){vapply(split(1:nrow(x),
                                            format(as.POSIXlt(x$Start),
                                                   "%Y-%m-%d %H:00:00",
                                                    tz='UTC')),
                                            length,
                                            0)})
visitHourAnimal <- unlist(visitHourAnimal)
visitHourAnimalDF <- read.table(text = names(visitHourAnimal), sep = ".")
visitHourAnimalDF$VisitCount <- as.numeric(visitHourAnimal)
names(visitHourAnimalDF) <- c("AnimalName", "Hour", "VisitCount")
visitHourAnimalDF$Hour <- as.POSIXct(visitHourAnimalDF$Hour)
```

```{r, echo = FALSE, fig.width = 12, fig.height = 6}
visitLines <- ggplot(visitHourAnimalDF,
                     aes(Hour, VisitCount, colour = AnimalName))
visitCountPlot <- visitLines + geom_line()
print(visitCountPlot)
```

```{r, echo = FALSE}
#create a data frame for visit per hour
hours <- vapply(split(1:nrow(visits),
                      format(as.POSIXlt(visits$Start),"%Y-%m-%d%H:00:00",tz='UTC')),
                      length,
                      0)
visitsPerHour <- as.numeric(hours)
hour <- as.POSIXct(names(hours))
visitCount <- data.frame(hour,visitsPerHour)
#Plot a line plot for the visitCount data frame
visitCountLine <- ggplot(visitCount, aes(hour, visitsPerHour)) +
                  geom_line() +
                  ylab("Count") +
                  xlab("") +
                  ggtitle("Visit count per hour")
```

This is the hour with the maximun number of visits  
```{r, echo = FALSE}
print(visitCount$hour[which.max(visitCount$visitsPerHour)])
```


This is the nosepokes count with the appropriate plot 
```{r, echo = FALSE}
#Plots nosepoke count per animal
nosepokesCountPlot <- ggplot(nosepokes, aes(AnimalName)) +
                      geom_bar() +
                      xlab("") +
                      ggtitle("Nosepokes Count")
print(table(nosepokes$AnimalName))
plot(nosepokesCountPlot)
```


Nosepokes over time for each animal
```{r, fig.width=10, fig.height=4, echo = FALSE}
#Plot Nosepokes over time using the ggplot2 package
nosepokePoints <- geom_point(aes(x = as.POSIXct(nosepokes$Start),
                                 y = nosepokes$AnimalName,
                                 colour = nosepokes$AnimalName),
                                 alpha = 0.15)
nosepokesPointsPlot <- ggplot() +
                       nosepokePoints +
                       xlab("") +
                       ylab("") +
                       ggtitle("Nosepokes") +
                       theme(legend.title=element_blank())
plot(nosepokesPointsPlot)
```

Check that drinking occoured only in the 19:00 23:59 time window, it shows the nosepokes where atleast one lick had happend, horizontal lines show the time window were the corner is active.
```{r, echo = FALSE, fig.width=10, fig.height=4}
#Plot licks over time using the ggplot2 package
lickIdx <- (nosepokes$LickNumber > 0)
lickPoints <- geom_point(aes(x =as.POSIXct(nosepokes$Start[lickIdx]),
                                 y = nosepokes$AnimalName[lickIdx],
                                 colour = nosepokes$AnimalName[lickIdx]),
                                 alpha = 0.3)
lickPointsPlot <- ggplot() +
                       lickPoints +
                       xlab("") +
                       ylab("") +
                       ggtitle("Nosepokes with at least one lick") +
                       theme(legend.title=element_blank())
plot(lickPointsPlot)

```


This is a plot showing the amount of nosepokes in the time window per day per animal  
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#find all the nosepokes which occoured in the time windows
library(lubridate)
hourNosepokes <- hour(ymd_hms(as.POSIXct(nosepokes$Start)))
inds <- 1:max(which(hourNosepokes == 23)) #indices of nosepokes in days                                            #with window
defModInds <- visits$VisitID[visits$ModuleName == "Default"]
hourNosepokes <- hourNosepokes[inds]
windowNosepokes <- nosepokes
dayNosepokes <- day(ymd_hms(as.POSIXct(
                    windowNosepokes$Start)))
windowNosepokes$Day <- dayNosepokes
windowNosepokes$Hour <- hourNosepokes
startHour <- 19
endHour <- 23
windowNosepokes <- windowNosepokes[!(nosepokes$VisitID %in% defModInds),]
p1 <- ggplot(windowNosepokes, aes(Day, fill=AnimalName)) +
        geom_bar() +
        ggtitle("Inside the time window nosepoke count")
print(p1)
```

This is the count of nosepokes per day per animal in the time windows
```{r, echo = FALSE}
##Add a moduleName column to windowNosepokes
windowNosepokes$ReactTime <- unlist(windowNosepokes$ReactTime)
windowNosepokes$ModuleName <- lapply(windowNosepokes$VisitID,function(x){
    mName <- visits$ModuleName[visits$VisitID == x]
    mName
})
windowNosepokes$ModuleName <- as.factor(unlist(windowNosepokes$ModuleName))
#make a data frame with the count of nosepokes in the time window for each day for each animal
results <-  expand.grid(animals$AnimalName, 
                        unique(windowNosepokes$Day),
                        unique(windowNosepokes$ModuleName))

names(results) <- c("AnimalName", "Day", "ModuleName")

#Set Start and End colums in data as postixct
windowNosepokes$Start <- as.POSIXct(windowNosepokes$Start) 
windowNosepokes$End <- as.POSIXct(windowNosepokes$End)
visits$Start <- as.POSIXct(visits$Start)

#Create a histogram for reaction times
source("findReactTimes3.R")
windowNosepokes$ReactTime <- findReactTimes3(visits, windowNosepokes)
windowNosepokes$TimeError[windowNosepokes$ReactTime < 0 | windowNosepokes$ReactTime > 0.9] <- 1
source("sumTimeErrors.R")
source("sumNosepokes.R")
TimeError <- as.numeric()
Count <- as.numeric()
for(i in 1:nrow(results)){
        TimeError[i] <- sumTimeErrors(windowNosepokes, 
                                      results$AnimalName[i], 
                                      results$Day[i], 
                                      results$ModuleName[i])
        Count[i] <-  sumNosepokes(windowNosepokes, 
                                      results$AnimalName[i], 
                                      results$Day[i], 
                                      results$ModuleName[i])
        
}
results$Count <- Count
results$TimeError <- TimeError
```

This is a plot showing the fraction of errors per time window per module,
the number of the module corresponds the anticipation time in seconds.
```{r, echo = FALSE, fig.width=10, fig.height=4, warning = FALSE, message = FALSE}
p3 <- ggplot(results,
             aes(Day,
                 TimeError/Count,
                 colour = AnimalName)) +
        geom_line() +
        geom_point(alpha = 0.5) +
        facet_grid(. ~ ModuleName) +
        ggtitle("Error fraction per day")
        
        
print(p3)
results$ErrorFraction <- results$TimeError/results$Count
```

This is an histogram showing the reaction times(values under zero are nosepokes before the light cue) by Day by anticipation time, the salmon colored line is the overall median reaction time:
```{r, echo = FALSE, fig.width=10, fig.height=9}
p4 <- ggplot(windowNosepokes, aes(ReactTime)) +
        geom_histogram(binwidth = 0.1) +
        scale_x_continuous(breaks = seq(0,25,1)) +
        ggtitle("Reaction times histogram") +
        geom_vline(xintercept =  median(windowNosepokes$ReactTime),
                   colour = "salmon", linetype = 5) +
        facet_grid(AnimalName ~ ModuleName)
        
print(p4)
print(paste0("The median reaction time in seconds:"
             ,median(windowNosepokes$ReactTime)))
```


```{r, echo = FALSE, fig.width=27, fig.height=27}
p5 <- ggplot(windowNosepokes, aes(ReactTime)) +
        geom_histogram(binwidth = 0.1) +
        scale_x_continuous(breaks = seq(0,25,1)) +
        ggtitle("Reaction times histogram") +
        geom_vline(xintercept =  median(windowNosepokes$ReactTime),
                   colour = "salmon", linetype = 5, alpha = 0.2) +
        facet_grid(AnimalName ~ ModuleName + Day)
print(p5)
```

These are the summarized results
```{r, echo = FALSE, fig.width=6, fig.height=8}
#Add sound and injection factor columns
results$Injection[results$Day == 22 & (results$AnimalName =="Animal 4" | results$AnimalName =="Animal 6" | results$AnimalName == "Animal 13")] <- "CNO"
results$Injection[results$Day == 22 & (results$AnimalName =="Animal 2" | results$AnimalName =="Animal 5" | results$AnimalName == "Animal 9")] <- "Saline"
results$Injection[results$Day == 23 & (results$AnimalName =="Animal 2" | results$AnimalName =="Animal 5" | results$AnimalName == "Animal 9")] <- "CNO"
results$Injection[results$Day == 23 & (results$AnimalName =="Animal 4" | results$AnimalName =="Animal 6" | results$AnimalName == "Animal 13")] <- "Saline"
results$Sound[results$ModuleName == "Module 1" | results$ModuleName == "Module 2" | results$ModuleName == "Module 3"] <- "Off"
results$Sound[!(results$ModuleName == "Module 1" | results$ModuleName == "Module 2" | results$ModuleName == "Module 3")] <- "On"

#Create a summary of all the results
resSum <- expand.grid(unique(results$Injection),
                      unique(results$Sound))
names(resSum) <- c("Injection", "Sound")

#Calculate the mean of each condition
resSum$Mean[resSum$Injection == "Saline" & resSum$Sound == "Off"] <- mean(results$ErrorFraction[results$Injection == "Saline" & results$Sound == "Off"])
resSum$Mean[resSum$Injection == "Saline" & resSum$Sound == "On"] <- mean(results$ErrorFraction[results$Injection == "Saline" & results$Sound == "On"])
resSum$Mean[resSum$Injection == "CNO" & resSum$Sound == "Off"] <- mean(results$ErrorFraction[results$Injection == "CNO" & results$Sound == "Off"])
resSum$Mean[resSum$Injection == "CNO" & resSum$Sound == "On"] <- mean(results$ErrorFraction[results$Injection == "CNO" & results$Sound == "On"])

#Calculate the SEM of each condition
resSum$SEM[resSum$Injection == "Saline" & resSum$Sound == "Off"] <- sd(results$ErrorFraction[results$Injection == "Saline" & results$Sound == "Off"])/sqrt(length(results$ErrorFraction[results$Injection == "Saline" & results$Sound == "Off"]))
resSum$SEM[resSum$Injection == "Saline" & resSum$Sound == "On"] <- sd(results$ErrorFraction[results$Injection == "Saline" & results$Sound == "On"])/sqrt(length(results$ErrorFraction[results$Injection == "Saline" & results$Sound == "On"]))
resSum$SEM[resSum$Injection == "CNO" & resSum$Sound == "Off"] <- sd(results$ErrorFraction[results$Injection == "CNO" & results$Sound == "Off"])/sqrt(length(results$ErrorFraction[results$Injection == "CNO" & results$Sound == "Off"]))
resSum$SEM[resSum$Injection == "CNO" & resSum$Sound == "On"] <- sd(results$ErrorFraction[results$Injection == "CNO" & results$Sound == "On"])/sqrt(length(results$ErrorFraction[results$Injection == "CNO" & results$Sound == "On"]))

#reorder factor levels to make the graph in the logical order
resSum$Injection <- factor(resSum$Injection,levels(resSum$Injection)[c(2,1)])
resSum$Sound <- factor(resSum$Sound,levels(resSum$Sound)[c(2,1)])

#Create a grouped bar graph for the results summary
p6 <- ggplot(resSum, aes(x=Sound, y=Mean, fill=Injection)) + 
      geom_bar(position=position_dodge(), stat="identity") +
      geom_errorbar(aes(ymin=Mean-SEM, ymax=Mean+SEM),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
        ggtitle("Mean of error fraction with SEM")
print(results)
print(resSum)
print(p6)

```

```{r, echo = FALSE, fig.width=6, fig.height=8, warning = FALSE, message = FALSE}
p7 <- ggplot(results,
             aes(Day,
                 TimeError/Count,
                 colour = AnimalName)) +
        geom_point(alpha = 0.75) +
        facet_grid(ModuleName ~ Injection) +
        ggtitle("Error fraction per day")
print(p7)
```
##Conclusion  
